---
title: "Final Project"
author: "Yunqiu Yao, Cenai Zhang, Yutian Mu, Murrel Pereira"
date: "12/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(stringr)
library(forcats)
library(boot)
```

## Clean the data
```{r}
### clean the data
gh = read_excel("./GHProject_Dataset.xlsx") %>% 
  clean_names %>% 
  separate(admitdtm, c("admit_dow","date","year"), ",") %>% 
  mutate(date = trimws(date, "left")) %>% 
  separate(date,c("month","sdate")," ") %>% 
  mutate(month = as.numeric(match(month,month.name)),
         sdate = as.numeric(sdate),
         year = as.numeric(year),
         admit_date = paste(month, sdate, year, sep="-"),
         admit_date = as.Date(admit_date, "%m-%d-%Y")) %>%
  select(-c(sdate,year,month)) %>% 
  mutate(log_losday = log(losdays2))

### remove the duplicate visits
gh = gh[order(gh$admit_date),]
gh_unique <- gh[!duplicated(gh$patientid),]

### recode the levels of categorical variables
gh_tidy = gh_unique %>% 
  mutate(cindex = as.factor(ifelse(cindex == 0, "normal", ifelse(cindex <= 2, "mild", ifelse(cindex <= 4, "moderate", "severe")))),
         is30dayreadmit = as.factor(is30dayreadmit),
         icu_flag = as.factor(icu_flag)) %>% 
  select(log_losday, losdays2, patientid, visitid, admit_date, admit_dow, everything(), -loshours, -facilityname, -facilityzip, -postalcode)
```


### Descriptive Statistics
```{r}
gh_tidy$gender[gh_tidy$gender == "Male"] = 0
gh_tidy$gender[gh_tidy$gender == "Female"] = 1

par(mfrow = c(1, 2))
hist(gh_tidy$losdays2, main = "Length of Stay")
hist(gh_tidy$log_losday, main = "Log Length of Stay")

par(mfrow = c(2, 2))
barplot(table(gh_tidy$admit_dow), las = 2, main = "Day of Week Frequency")
barplot(table(gh_tidy$is30dayreadmit), las = 1, main = "Is patient 30 day readmit?")
barplot(table(gh_tidy$mews), las = 2, main = "Modified Early Warning Score (MEWS)")
barplot(table(gh_tidy$cindex), las = 2, main = "Charlson comorbidity index (CCI)")

par(mfrow = c(3, 2))
barplot(table(gh_tidy$icu_flag), main = "If patient visited ICU")
barplot(table(gh_tidy$gender), main = "Gender")
barplot(table(gh_tidy$race), las =2, main = "Race Frequency")
barplot(table(gh_tidy$religion), las = 2, main = "Religion Frequency")
barplot(table(gh_tidy$maritalstatus), las = 2, main = "Marital Status Frequency")
barplot(table(gh_tidy$insurancetype), las = 1, main = "Insurance Type")

par(mfrow = c(3, 3))
hist(gh_tidy$evisit, main = "ER visits in past 6 months")
hist(gh_tidy$ageyear, main ="Age Histogram") #might need a log transformation
hist(gh_tidy$bmi, main = "BMI")
hist(gh_tidy$bpsystolic, main = "BP Systolic") #120 - 140 mmHg normal range
hist(gh_tidy$bpdiastolic, main = "BP Diastolic") #80 -90 mmHg normal range
hist(gh_tidy$o2sat, main = "O2 Saturation") #Normal is  <100 - check for outliers
hist(gh_tidy$temperature, main = "Temperature") #Normal range is 36.1 C - 37.2 C - check for outliers; can go upto 42C
hist(gh_tidy$heartrate, main = "Heart Rate") #40 - 100 beats per minute
hist(gh_tidy$respirationrate, main = "Respiration Rate") #12 - 25 breaths per minute in normal adult

par(mfrow = c(3, 3))
boxplot(gh_tidy$evisit, main = "ER visits in past 6 months")
boxplot(gh_tidy$ageyear, main ="Age Histogram") #might need a log transformation
boxplot(gh_tidy$bmi, main = "BMI")
boxplot(gh_tidy$bpsystolic, main = "BP Systolic") #120 - 140 mmHg normal range
boxplot(gh_tidy$bpdiastolic, main = "BP Diastolic") #80 -90 mmHg normal range
boxplot(gh_tidy$o2sat, main = "O2 Saturation") #Normal is  <100 - check for outliers
boxplot(gh_tidy$temperature, main = "Temperature") #Normal range is 36.1 C - 37.2 C - check for outliers; can go upto 42C
boxplot(gh_tidy$heartrate, main = "Heart Rate") #40 - 100 beats per minute
boxplot(gh_tidy$respirationrate, main = "Respiration Rate") #12 - 25 breaths per minute in normal adult

#Based on Boxplot for Temperature, many outliers are identified. NA anything outside the minimum and maximum 1.5IQR
#1.5IQR; if above 111.93 or below 46.89, remove from data set; This range is more realistic for an adult
gh_tidy$temperature[gh_tidy$temperature < 36.37] <- NA
gh_tidy$temperature[gh_tidy$temperature > 37.12] <- NA

#Based on Boxplot for Temperature, many outliers are identified. NA anything outside the minimum and maximum 1.5IQR
#1.5IQR; if above 111.93 or below 46.89, remove from data set; This range is more realistic for an adult
gh_tidy$o2sat[gh_tidy$o2sat < 94.43] <- NA
gh_tidy$o2sat[gh_tidy$o2sat > 100.61] <- NA


#Based on Boxplot for Heart Rate, many outliers are identified. NA anything outside the minimum and maximum 1.5IQR
#1.5IQR; if above 111.93 or below 46.89, remove from data set; This range is more realistic for an adult
gh_tidy$heartrate[gh_tidy$heartrate < 46.89] <- NA
gh_tidy$heartrate[gh_tidy$heartrate > 111.93] <- NA

#Based on Boxplot for Respiration Rate, many outliers are identified. NA anything outside the minimum and maximum 1.5IQR
#1.5IQR; if above 20.48 or below 15.085, remove from data set; This range is more realistic for an adult
gh_tidy$respirationrate[gh_tidy$respirationrate < 15.085] <- NA
gh_tidy$respirationrate[gh_tidy$respirationrate > 20.48] <- NA

#Based on Boxplot for BP Diastolic, many outliers are identified. NA anything outside the minimum and maximum 1.5IQR
#1.5IQR; if above 89.56 or below 54.19, remove from data set; This range is more realistic for an adult
gh_tidy$bpdiastolic[gh_tidy$bpdiastolic < 54.19] <- NA
gh_tidy$bpdiastolic[gh_tidy$bpdiastolic > 89.56] <- NA

#Based on Boxplot for BP Systolic, many outliers are identified. NA anything outside the minimum and maximum 1.5IQR
#1.5IQR; if above 164.86 or below 94.57, remove from data set; This range is more realistic for an adult
gh_tidy$bpsystolic[gh_tidy$bpsystolic < 94.57] <- NA
gh_tidy$bpsystolic[gh_tidy$bpsystolic > 164.86] <- NA

par(mfrow = c(3, 3))
hist(gh_tidy$evisit, main = "ER visits in past 6 months")
hist(gh_tidy$ageyear, main ="Age Histogram") #might need a log transformation
hist(gh_tidy$bmi, main = "BMI")
hist(gh_tidy$bpsystolic, main = "BP Systolic") #120 - 140 mmHg normal range
hist(gh_tidy$bpdiastolic, main = "BP Diastolic") #80 -90 mmHg normal range
hist(gh_tidy$o2sat, main = "O2 Saturation") #Normal is  <100 - check for outliers
hist(gh_tidy$temperature, main = "Temperature") #Normal range is 36.1 C - 37.2 C - check for outliers; can go upto 42C
hist(gh_tidy$heartrate, main = "Heart Rate") #40 - 100 beats per minute
hist(gh_tidy$respirationrate, main = "Respiration Rate") #12 - 25 breaths per minute in normal adult

par(mfrow = c(3, 3))
boxplot(gh_tidy$evisit, main = "ER visits in past 6 months")
boxplot(gh_tidy$ageyear, main ="Age Histogram") #might need a log transformation
boxplot(gh_tidy$bmi, main = "BMI")
boxplot(gh_tidy$bpsystolic, main = "BP Systolic") #120 - 140 mmHg normal range
boxplot(gh_tidy$bpdiastolic, main = "BP Diastolic") #80 -90 mmHg normal range
boxplot(gh_tidy$o2sat, main = "O2 Saturation") #Normal is  <100 - check for outliers
boxplot(gh_tidy$temperature, main = "Temperature") #Normal range is 36.1 C - 37.2 C - check for outliers; can go upto 42C
boxplot(gh_tidy$heartrate, main = "Heart Rate") #40 - 100 beats per minute
boxplot(gh_tidy$respirationrate, main = "Respiration Rate") #12 - 25 breaths per minute in normal adult


##Late Weekday: Thursday - Sunday, 0
##Early Weekday: Monday - Wednesday, 1

gh_tidy$admit_dow[gh_tidy$admit_dow == "Thursday"] <- 0
gh_tidy$admit_dow[gh_tidy$admit_dow == "Friday"] <- 0
gh_tidy$admit_dow[gh_tidy$admit_dow == "Saturday"] <- 0
gh_tidy$admit_dow[gh_tidy$admit_dow == "Sunday"] <- 0
gh_tidy$admit_dow[gh_tidy$admit_dow == "Monday"] <- 1
gh_tidy$admit_dow[gh_tidy$admit_dow == "Tuesday"] <- 1
gh_tidy$admit_dow[gh_tidy$admit_dow == "Wednesday"] <- 1


class(gh_tidy$admit_dow) <- "binary"

barplot(table(gh_tidy$admit_dow), las = 2, main = "Day of Week Frequency")


```

## Select the variables and fit the linear model
```{r}
gh_reg = gh_tidy %>%
  select(-losdays2, -patientid, -visitid, -admit_date, -bmi, -mews) %>%
  na.omit() %>% 
  mutate(cindex = fct_relevel(cindex, "normal")) 

gh_reg_icu1 = gh_reg %>% filter(icu_flag == 1) %>% select(-icu_flag)
gh_reg_icu0 = gh_reg %>% filter(icu_flag == 0) %>% select(-icu_flag)

#fit models
model_icu1 <- lm(gh_reg_icu1$log_losday~ ., data = gh_reg_icu1)
summary(model_icu1)
anova(model_icu1)

model_icu0 <- lm(log_losday~ ., data = gh_reg_icu0)
summary(model_icu0)
anova(model_icu0)
```


```{r}
summary_functn <- function(x){
  list(variable = colnames(x),
       minimum = apply(x, 2, min),
       median = apply(x, 2, median),
       mean = apply(x, 2, mean),
       maximum = apply(x, 2, max),
       sd = apply(x, 2, sd)) %>%
    as.tibble()
}

gh_reg %>%
select(log_losday,evisit,ageyear,bpsystolic,o2sat,temperature,heartrate,respirationrate,bpdiastolic) %>%
summary_functn()

gh_categorical = gh_reg %>%
mutate(admit_dow=as.factor(admit_dow),
is30dayreadmit=as.factor(is30dayreadmit),
gender=as.factor(gender),
race=as.factor(race),
religion=as.factor(religion),
maritalstatus=as.factor(maritalstatus),
insurancetype=as.factor(insurancetype)) %>%
select(admit_dow,is30dayreadmit,cindex,icu_flag,gender,race,religion,maritalstatus,insurancetype)
summary(gh_categorical)
```

## Model fitting: Backward, Stepwise, and Forward
```{r}
# forward
forward_icu0 = step(model_icu0, direction = 'forward')
summary(forward_icu0)
anova(forward_icu0)

forward_icu1 = step(model_icu1, direction = 'forward')
summary(forward_icu1)



summary(lm(formula = gh_reg$log_losday ~ admit_dow + is30dayreadmit + mews + cindex + evisit + icu_flag + ageyear + gender + race + religion + maritalstatus + insurancetype, data = gh_reg))
```
Forward keeps all of the original predictors, but most of them are not significant.

```{r model_building}
# for ICU_1
model_icu1 <- lm(gh_reg_icu1$log_losday~ ., data = gh_reg_icu1)
anova(model_icu1)
step1 <- update(model_icu1 , . ~ . -bpdiastolic)
anova(step1)
step2 <- update(step1, . ~ . -insurancetype)
anova(step2)
step3 <- update(step2, . ~ . -gender)
anova(step3)
step4 <- update(step3, . ~ . -bpsystolic )
anova(step4)
step5 <- update(step4, . ~ . -heartrate)
anova(step5)
step6 <- update(step5, . ~ . -respirationrate)
anova(step6)
step7 <- update(step6, . ~ . -race)
anova(step7)
step8 <- update(step7, . ~ . -religion)
anova(step8)
summary(step8)

# for ICU_0
model_icu0 <- lm(gh_reg_icu0$log_losday~ ., data = gh_reg_icu0)
anova(model_icu0)
step1 <- update(model_icu0 , . ~ . -race)
anova(step1)
step2 <- update(step1, . ~ . -religion)
anova(step2)
step3 <- update(step2, . ~ . -gender )
anova(step3)
summary(step3)
```


```{r}
# For stepwise
stepwise_icu0 = step(model_icu0, direction = 'both')
summary(stepwise_icu0)
anova(stepwise_icu0)

stepwise_icu1 = step(model_icu1, direction = 'both')
summary(stepwise_icu1)
anova(stepwise_icu1)
```


```{r}
# both and backward are same.
#is30dayreadmit1+mews(*3)+cindex(*3)+evisit+ageyear+insurance(*2)
```
The result of both and backward are the same, and we can see in the summary most of the predictors are significant. We'll this model for the further analysis.

```{r}
#criterion-based 
best <- function(model, ...) 
{
  subsets <- leaps::regsubsets(formula(model), model.frame(model), ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  

best(model_icu0, nbest=1)
```

## Model diagnosis
```{r diagnosis}
fit = lm(log_losday ~ is30dayreadmit + mews + cindex + evisit + ageyear + insurancetype, data=gh_reg)
par(mfrow = c(2, 2))
plot(fit)
```

## Model Validation
### LOOCV
```{r validation}
#LOOCV--unbiased estimate of test error
glm.fit<-glm(log_losday ~ is30dayreadmit + mews + cindex + evisit + ageyear + insurancetype, data=gh_reg)
cv.err<-cv.glm(gh_reg,glm.fit)
cv.err$delta   

```
Delta value is the average mean-squared error. According to the lecture, the two delta values should be similar. We use the first one, and the second value is bias corrected.


### Bootstrap
```{r bootstrap}
set.seed(1)
gh_reg_df = as.data.frame(gh_reg)
attach(gh_reg_df)

# Our usual regression, no bootstrap yet
boot.fn<-function(data, index){
	return(coef(lm(log_losday ~ is30dayreadmit + mews + cindex + evisit + ageyear + insurancetype, data=gh_reg_df,subset=index)))
}

# Use function boot() to repeat the sampling 10000 times.
# Repeat 10000 times to get the estimates, SEs ad bias
boot(gh_reg_df, boot.fn, 100)
```
The results are the standard errors for the estimated coefficient, and it should be around zero. Our results show that the se of all the predictors are quite small. 


